Two-Dimensional Vesicle Hydrodynamics from Hydrophobic Attraction Potential

### 2021-09-28 doc/main_jfm.tex
final text changes

To reach consistent simulation outcomes, we first solve for a baseline
JP vesicles that is suitably close to equilibrium. We start with an
assumed configuration of $N=58$ JP in the form of two, circular,
apposing monolayers of about $8$ nm in radius. The translational and
rotational velocities vanish exponentially with an approximate decay
rate $\sim 4.6$ $\upmu$s$^{-1}$. An equilibrium configuration is
therefore rapidly attained. This equilibrium configuration serves as the
initial data in the subsequent background flow simulations.

@@

\begin{figure}
\begin{center}
  \includegraphics[width=11cm]{Figure5_Wrapper.pdf}\\
  \includegraphics[width=11cm]{Figure5B_Wrapper.pdf}
%\includegraphics[width=0.5\textwidth]{Slip.eps}\\
%\includegraphics[width=0.4\textwidth]{Slip2.eps}
\end{center} 
\caption{\label{figure5} There is inter-monolayer slip present in the JP
  vesicle tank-treading motion at all shear rates. Panel (a) tracks a
  pair of particles (blue and yellow). The shear rate is $\chi=0.005$
  and in the right figure, the yellow particle will complete two and a
  quarter revolutions in the same time the blue particle completes two
  revolutions. The curves in panel (b) are the distance the outer
  leaflet has slid past the inner leaflet. The slopes of the curves
  gives the slip velocity. With the exception of $\chi = 0.0035$, the
  slip velocity is generally monotonic in the shear rate.}
\end{figure}

@@

We point out that the range of values for $\chi$ where we measured for
tank-treading correspond to shear rates $\dot \gamma = \mathcal{O}(10^6$
s$^{-1}$) which gives fluid velocities $\mathcal{O}($m s$^{-1}$$)$ in
the vicinity of the vesicle. While large, these values are identical to
ones used in prior MD studies (\cite{Brandner2019}) and are a
consequence of the fact that larger shear rates are required to produce
the viscous stresses needed to appreciably deform smaller vesicles.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Inter-Monolayer Friction}
We observe inter-monolayer slip in the tank-treading, JP vesicle at all
shear rates. Since they are not bound, the two leaflets of the vesicle
are able to slide past one another. This sliding is limited by viscous
friction of the aqueous gaps between particles and by the constant
unbinding and binding of particles pairs in apposing leaflets. The
phenomenon is consistent with prior MD results (\cite{Zgorski2019,
denOtter2007}).

Figure~\ref{figure5}a illustrates the process by tracking the distances
traveled by a pair of particles. In the left image, the blue and yellow
particle lie next to each other. In the right-most panel, the yellow
particle has traveled farther than the blue particle. This suggests that
the tangential velocity of the outer leaflet is larger than that of the
inner leaflet. The difference in their velocities is the slip velocity
$U(t)$ and $s(t) = \int_0^t dU $ records the distance one leaflet has
slid past the other.

Figure~\ref{figure5}b plots the sliding distance for various shear
rates. From the data, we define an inter-monolayer friction coefficient 
\begin{align}
  b = \frac{\langle F\rangle}{\langle L \rangle \langle U\rangle},
\end{align}
where $F$ is the tangential force jump, $L$ is the length of the
midplane $C_M$ (Figure~\ref{figure5}a), $U$ is the slip velocity, and
$\langle \cdot \rangle$ denotes the time average. We define the
tangential force jump $F$ to be the tangential shear force on the outer
leaflet minus the tangential shear force on the inner leaflet. However,
it does not make sense to integrate along a curve passing through the
particles.  To calculate the tangential force jump $F$, we instead let 
\begin{align}
F_h = \int_{\mathcal{C}_M^+} 
  \ttau \cdot \sigma \cdot \nnu \,\dif s - \int_{\mathcal{C}_M^-} 
 \ttau \cdot \sigma \cdot \nnu \,\dif s 
\end{align}
where $C_M^+$ and $C_M^-$ are target curves obtained by moving the
midplane curve $C_M$ (Figure~\ref{figure5}a) a distance $h$ outward,
respectively inward, along the unit normal. We sample $F_h$ for $h = \pm
1.5, \pm 2.5, \pm 3.5$ times the particle radius and define $F$ by
extrapolating to zero distance.

\begin{table}
\caption{Friction Coefficients}
\centering
\begin{tabular}{c c c c c c c c }
 $\dot\gamma$  (ns$^{-1}$) & 0.0020   &  0.0025 &  0.0030 &  0.0035 &  0.0040 & 0.0045 & 0.0050  \\
\hline                    
%$\dot\gamma$ (ns$^{-1}$)        & 0.0033 & 0.0041 & 0.0049 & 0.0057 & 0.0066 & 0.0074 & 0.0082\\
%$F_{f}$ (pN nm$^{-1}$)           & 1.84 & 4.90 & 11.36 & 8.16 & 6.14 &  17.69 &  24.33 \\[1ex]
%$\dot\theta$ (nm ns$^{-1}$)           & $6.84\times10^{-5}$ & $1.31\times10^{-4}$ & $3.46\times10^{-4}$ & $2.32\times10^{-4}$ & $1.81\times10^{-4}$ & $4.91\times10^{-4}$ & $5.83\times10^{-4}$\\ [1ex]
%$b$ (pN$\cdot$ns nm$^{-3}$)    & 0.22 & 0.87*  & 0.22 &0.21 & 1.29* & 0.31 & 0.30\\ 
$b$ (pN$\cdot$ns nm$^{-3}$)    & 1.01 & 3.40 &   3.70  &   5.01 &   5.38  &   5.75  &   6.25 \\ 
\hline    
\end{tabular} 
\label{table1}
\end{table}

Table~\ref{table1} provides dimensional friction coefficients $b$ for
several shear rates. The coefficients generally lie within the range 1
to 6 pN ns nm$^{-3}$ which is in good quantitative agreement with values
previously reported in the literature (\cite{sch-vla-mik2010}
and~\cite{denOtter2007}). \cite{denOtter2007} give $b = 2.4 \times 10^6$
Pa m$^{-1}$ s = 2.4 pN nm$^{-3}$ ns from their MD study, for example.

@@

\subsection{Stretching and Permeability}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{Figure2_Wrapper.pdf}
%\includegraphics[height=2in]{relax2.eps}
%\includegraphics[height=2in]{relax.eps}\\
%\includegraphics[height=2.in]{N58_0pres.pdf}
%\includegraphics[height=2.in]{N58_5000pres.pdf}
%  \todo[inline]{Put pressure of each of the two configurations below
%  this.}
\caption{\label{figure2} In JP vesicles, tension and pressure relax as
  fluid is expelled through the particle interstices. In panel (a), the
  thin curve plots the mean of the norm of the particle translational
  velocities. The thick curve plots the mean of the absolute value of
  the particle angular velocities. At $t = 0$ $\upmu$s, the vesicle has
  a spatially constant internal pressure 0.05~pN nm$^{-1}$ (red color,
  panel (b)). In panel (c), pressure has dissipated and the vesicle has
  moved from its initial configuration (gray circles), to the
  equilibrium configuration (black circles). The data in panels (d) and
  (e) support the linear relationship expressed by
  \eqref{eq:LaplacePressure} and \eqref{eq:perm}. Circles are for a
  vesicle with 17 nm initial radius; square are for a 34 nm initial
  radius. Asterisks are for a 18 nm initial radius but with smaller,
  1.25 nm diameter particles. In panel (f), the theoretical time course
  \eqref{eq:perm_solution} overlaps the data using constants $K_A$ and
  $P$ derived from panels (d) and (e). }
\end{figure}
Finally, we show that the JP vesicle behaves as an extensible, permeable
membrane. Figures~\ref{figure2}b shows an initial, non-equilibrium JP
vesicle suspended in a quiescent flow $\uu_{\infty} = 0$. The color map
plots the pressure $p$. The red color in Figures~\ref{figure2}b shows a
constant, positive internal pressure (0.05~pN nm$^{-1}$) and the white
shows a constant zero, external pressure. There is some fluid flow and
the pressure vanishes as the configuration tends toward the equilibrium
state (Figures~\ref{figure2}c).

What could be the source of this drop in pressure? In continuum membrane
mechanics, changes in surface area give rise to a membrane tension
$\tilde \gamma = K_A(A/A_0 - 1)$ where $A$ is the membrane surface area
and $A_0$ is the reference surface area. The area modulus $K_A$ of
bilayers is about $240$ pN nm$^{-1}$ (\cite{NaTr00}). In the
two-dimensional vesicles, the tension becomes
\begin{align}
\label{eq:stretch}
\tilde \gamma = K_A\left(\frac{L}{L_0} - 1 \right),
\end{align}
where $L$ and $L_0$ are the vesicle arc length and resting length,
respectively. Moreover, stretched, circular vesicles has a Laplace pressure 
\begin{align}
\label{eq:LaplacePressure}
\Delta p = \frac{\tilde \gamma}{R},
\end{align}
where $\Delta p$ is the difference in internal pressure to the pressure
at infinity and $R^{-1}$ is the total curvature of the circular
cylinder.

Figure~\ref{figure2}c plots the data for the pressure jump $\Delta p$
between the particle center and the far-field (see
Figures~\ref{figure2}c and~\ref{figure2}d). We use $R = L/(2\pi)$ for
the vesicle radius, and the horizontal axis is the relative stretch.
The circles are data for a vesicle with radius 17 nm. The linear fit
(solid lines) shows that $R\Delta P$ is proportional to the relative
stretching $L/L_0 - 1$. The squares are for a vesicle with twice the
radius 34 nm, and the data overlap supports that the proportionality
constant $K_A$ is a stretching modulus that is independent of vesicle
size. The data give $K_A = 170 \pm 9$ pN nm$^{-1}$ which is good
agreement with the area moduli of lipid bilayer as determined by
experiment. To evaluate how changes to particle size lead to different
physical properties, the asterisk symbols are for an 18 nm radius
vesicle consisting of particles with diameter 1.25 nm, instead of the
usual 2.5 nm. These data give smaller stretching modulus of $K_A =
102$~pN~nm$^{-1}$.

The particles in our setup do not abut but rather have small gaps due to
repulsive forces. The gaps allow for some fluid flux across the JP
bilayer, which can be quantified by the equation
\begin{align}
  \label{eq:perm} 
  \frac{dA}{dt} = -P L \Delta p,
\end{align}
where $P$ is a hydraulic permeability constant (\cite{chabanon2017,
qua-gan-you2021}). Figure~\ref{figure2}e shows that the data for
$L\Delta p$ and $dA/dt$ obey the linear relationship expressed by
\eqref{eq:perm}. The slopes of the linear fits give the hydraulic
permeabilities $P = 0.0296$ nm$^3$ ns$^{-1}$ pN$^{-1}$ and $P = 0.0283$
nm$^3$ ns$^{-1}$ pN$^{-1}$ for the 17 nm and 34 nm radius cases,
respectively. Like the area modulus, permeability is independent of
vesicle size.

@@

Since the vesicles in Figure~\ref{figure2}a are nearly circular, we can
combine~\eqref{eq:stretch},~\eqref{eq:LaplacePressure},
and~\eqref{eq:perm}, to derive
\begin{align}
\label{eq:perm_solution}
L_0(L-L(0)) + L_0^2 \ln\left|\frac{L-L_0}{L(0)-L_0}\right| = -4\pi^2 P K_A t.
\end{align}
The theoretical time courses for \eqref{eq:perm_solution} are in good
quantitative agreement with the JP data (Figure~\ref{figure2}f).

>>>>

To reach consistent simulation outcomes, we first solve for a baseline
JP vesicles that is suitably close to equilibrium. We start with an
assumed configuration of $N=58$ JP in the form of two, circular,
apposing monolayers of about $8$ nm in radius. The norms of the translational and
rotational velocities vanish exponentially with an approximate decay
rate $\sim 4.6$ $\upmu$s$^{-1}$. An equilibrium configuration is
therefore rapidly attained. This equilibrium configuration serves as the
initial data in the subsequent background flow simulations.

@@

\begin{figure}
\begin{center}
  \includegraphics[width=11cm]{Figure5_Wrapper.pdf}\\
  \includegraphics[width=11cm]{Figure5B_Wrapper.pdf}
%\includegraphics[width=0.5\textwidth]{Slip.eps}\\
%\includegraphics[width=0.4\textwidth]{Slip2.eps}
\end{center} 
\caption{\label{figure5} Inter-monolayer slip is present in the JP
  vesicle tank-treading motion at all shear rates. Panel (a) tracks a
  pair of particles (blue and yellow). The shear rate is $\chi=0.005$
  and in the right figure, the yellow particle will complete two and a
  quarter revolutions in the same time the blue particle completes two
  revolutions. The curves in panel (b) are the distance the outer
  leaflet has slid past the inner leaflet. The slopes of the curves
  give the slip velocity. With the exception of $\chi = 0.0035$, the
  slip velocity is generally monotonic in the shear rate.}
\end{figure}

@@

We point out that the range of values for $\chi$ where we measured for
tank-treading correspond to shear rates $\dot \gamma = \mathcal{O}(10^6$
s$^{-1}$) which gives fluid velocities $\mathcal{O}($m s$^{-1}$$)$ in
the vicinity of the vesicle. While large, these orders of magnitude are identical to
ones used in prior MD studies (\cite{Brandner2019}) and are a
consequence of the fact that larger shear rates are required to produce
the viscous stresses needed to appreciably deform smaller vesicles.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Inter-Monolayer Friction}
We observe inter-monolayer slip in the tank-treading, JP vesicle at all
shear rates. Since they are not bound, the two leaflets of the vesicle
are able to slide past one another. This sliding is limited by viscous
friction of the aqueous gaps between particles and by the constant
unbinding and binding of particles pairs in apposing leaflets. 
Previous studies \cite{Zgorski2019, denOtter2007}
have considered inter-monolayer slip in lipid bilayers simulated by MD.

Figure~\ref{figure5}a illustrates inter-monolayer slip by tracking the distances
traveled by a pair of particles along the midplane curve. In the left image, the blue and yellow
particle lie next to each other. In the right-most panel, the yellow
particle has traveled farther than the blue particle. This suggests that
the outer tangential velocity, obtained by projecting the velocity of the
outer leaflet onto the midplane curve, 
is larger than that of the inner leaflet.

From the data, we obtain an inter-monolayer friction coefficient 
\begin{align}
  b =  \frac{\langle F \rangle}{\langle L   U \rangle} ,
\end{align}
where $F$ is the tangential force jump, $L$ is the length of the
midplane $C_M$ (Figure~\ref{figure5}a), and $U$ is the slip velocity.
The time average $\langle \cdot \rangle$ is necessary to avoid division by zero
whenever slip velocity vanishes.

\begin{table}
\caption{Friction Coefficients}
\centering
\begin{tabular}{c c c c c c c c }
 $\dot\gamma$  (ns$^{-1}$) & 0.0020   &  0.0025 &  0.0030 &  0.0035 &  0.0040 & 0.0045 & 0.0050  \\
\hline                    
%$\dot\gamma$ (ns$^{-1}$)        & 0.0033 & 0.0041 & 0.0049 & 0.0057 & 0.0066 & 0.0074 & 0.0082\\
%$F_{f}$ (pN nm$^{-1}$)           & 1.84 & 4.90 & 11.36 & 8.16 & 6.14 &  17.69 &  24.33 \\[1ex]
%$\dot\theta$ (nm ns$^{-1}$)           & $6.84\times10^{-5}$ & $1.31\times10^{-4}$ & $3.46\times10^{-4}$ & $2.32\times10^{-4}$ & $1.81\times10^{-4}$ & $4.91\times10^{-4}$ & $5.83\times10^{-4}$\\ [1ex]
%$b$ (pN$\cdot$ns nm$^{-3}$)    & 0.22 & 0.87*  & 0.22 &0.21 & 1.29* & 0.31 & 0.30\\ 
$b$ (pN ns nm$^{-3}$)    & 0.43 & 1.19 &   0.42  & 0.69 &   0.97  &   0.80  &  1.07 \\ 
\hline    
\end{tabular} 
\label{table1}
\end{table}

The tangential force jump $F$ equals the 
tangential shear force on the outer
leaflet minus the tangential shear force on the inner leaflet. 
To calculate  $F$, we first let 
\begin{align}
F_h = \int_{\mathcal{C}_M^+} 
  \ttau \cdot \sigma \cdot \nnu \,\dif s - \int_{\mathcal{C}_M^-} 
 \ttau \cdot \sigma \cdot \nnu \,\dif s 
\end{align}
where $C_M^+$ and $C_M^-$ are target curves obtained by projecting the
midplane curve $C_M$  a distance $h$ outward,
respectively inward, along its unit normal vector field (Figure~\ref{figure5}a). We sample $F_h$ for 
$h = \pm 1.3, \pm 1.5,  \pm 1.7, \pm 1.9$ times the particle radius and define $F$ by
extrapolating to zero distance. This avoids integrating along a curve passing directly through the
particles.

To calculate $U$, we let $v_+(\mathbf{x})$ and $v_-(\mathbf{x})$ 
be the tangential velocity of the outer, respectively inner, leaflet.  Then 
\begin{equation}
U = \frac{1}{L}\int_{C_M}v_+(\mathbf{x})(1 - \delta \kappa) - v_-(\mathbf{x})(1 +  \delta \kappa) \, ds,
\end{equation}
where $ \delta $ is the distance from the leaflet centers to the midplane and $\kappa$ is the curvature. 
The factors $\pm  \delta \kappa$ are needed to project the leaflet velocities, defined on the particle centers,
onto the midplane curve. 
Finally, the function $s(t) = \int_0^t dU $ records the distance one leaflet has
slid past the other. Figure~\ref{figure5}b plots the sliding distance for various shear
rates.

Table~\ref{table1} provides dimensional friction coefficients $b$ for
several shear rates. 
The coefficients are 0.79 $\pm$ 0.3 pN ns nm$^{-3}$ which is in good quantitative agreement with values
previously reported in the literature (\cite{sch-vla-mik2010}
and~\cite{denOtter2007}). \cite{denOtter2007} give $b = 2.4 \times 10^6$
Pa m$^{-1}$ s = 2.4 pN ns nm$^{-3}$  from their MD study, for example.

@@

\subsection{Stretching and Permeability}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{Figure2_Wrapper.pdf}
%\includegraphics[height=2in]{relax2.eps}
%\includegraphics[height=2in]{relax.eps}\\
%\includegraphics[height=2.in]{N58_0pres.pdf}
%\includegraphics[height=2.in]{N58_5000pres.pdf}
%  \todo[inline]{Put pressure of each of the two configurations below
%  this.}
\caption{\label{figure2} In JP vesicles, tension and pressure relax as
  fluid is expelled through the particle interstices. In panel (a), the
  thin curve plots the mean of the norm of the particle translational
  velocities. The thick curve plots the mean of the absolute value of
  the particle angular velocities. At $t = 0$ $\upmu$s, the vesicle has
  a spatially constant internal pressure 0.05~pN nm$^{-2}$ (red color,
  panel (b)). In panel (c), pressure has dissipated and the vesicle has
  moved from its initial configuration (gray circles), to the
  equilibrium configuration (black circles). The data in panels (d) and
  (e) support the linear relationship expressed by
  \eqref{eq:LaplacePressure} and \eqref{eq:perm}. Circles are for a
  vesicle with 17 nm initial radius; square are for a 34 nm initial
  radius. Asterisks are for a 18 nm initial radius but with smaller,
  1.25 nm diameter particles. In panel (f), the theoretical time course
  \eqref{eq:perm_solution} overlaps the data using constants $K_A$ and
  $P$ derived from panels (d) and (e). }
\end{figure}
Finally, we show that the JP vesicle behaves as an extensible, permeable
membrane. Figures~\ref{figure2}b shows an initial, non-equilibrium JP
vesicle suspended in a quiescent flow $\uu_{\infty} = 0$. The color map
plots the pressure $p$. The red color in Figures~\ref{figure2}b shows a
spatially constant, positive internal pressure (0.05~pN nm$^{-2}$) and the white
shows a spatially constant zero, external pressure. There is some fluid flow and
the pressure vanishes as the configuration tends toward the equilibrium
state (Figures~\ref{figure2}c).

What could be the source of this drop in pressure? In membrane
continuum  mechanics, changes in surface area give rise to a membrane tension
$\tilde \gamma = K_A(A/A_0 - 1)$ where $A$ is the membrane surface area
and $A_0$ is the reference surface area. The area modulus $K_A$ of
bilayers is about $240$ pN nm$^{-1}$ (\cite{NaTr00}). In the
two-dimensional vesicles, the tension becomes
\begin{align}
\label{eq:stretch}
\tilde \gamma = K_A\left(\frac{L}{L_0} - 1 \right),
\end{align}
where $L$ and $L_0$ are the vesicle arc length and resting length,
respectively. Moreover, stretched, circular vesicles has a Laplace pressure 
\begin{align}
\label{eq:LaplacePressure}
\Delta p = \frac{\tilde \gamma}{R},
\end{align}
where $\Delta p$ is the difference in internal pressure to the pressure
at infinity and $R^{-1}$ is the total curvature of the circular
cylinder.

Figure~\ref{figure2}c plots the data for the pressure jump $\Delta p$
between the particle center and the far-field (see
Figures~\ref{figure2}c and~\ref{figure2}d). We use $R = L/(2\pi)$ for
the vesicle radius, and the horizontal axis is the relative stretch.
The circles are data for a vesicle with radius 17 nm. The linear fit
(solid lines) shows that $R\Delta P$ is proportional to the relative
stretching $L/L_0 - 1$. The squares are for a vesicle with twice the
radius 34 nm, and the data overlap supports that the proportionality
constant $K_A$ is a stretching modulus that is independent of vesicle
size. The data give $K_A = 170 \pm 9$ pN nm$^{-1}$ which is in good
agreement with the experimentally obtained area moduli of lipid bilayer. 
To evaluate how changes to particle size lead to different
physical properties, the asterisk symbols are for an 18 nm radius
vesicle consisting of particles with diameter 1.25 nm, instead of the
usual 2.5 nm. These data give a smaller stretching modulus of $K_A =
102$~pN~nm$^{-1}$.

The particles in our setup do not abut but rather have small gaps due to
repulsive forces. The gaps allow for some fluid flux across the JP
bilayer, and in membrane mechanics aqueous flux is quantified by the equation
\begin{align}
  \label{eq:perm} 
  \frac{dA}{dt} = -P L \Delta p,
\end{align}
where $P$ is a hydraulic permeability constant (\cite{chabanon2017,
qua-gan-you2021}). Figure~\ref{figure2}e shows that the data for
$L\Delta p$ and $dA/dt$ obey the linear relationship expressed by
\eqref{eq:perm}. The slopes of the linear fits give the hydraulic
permeabilities $P = 0.0296$ nm$^3$ ns$^{-1}$ pN$^{-1}$ and $P = 0.0283$
nm$^3$ ns$^{-1}$ pN$^{-1}$ for the 17 nm and 34 nm radius cases,
respectively. Like the area modulus, the data give a permeability 
that is independent of vesicle size.

@@

Since the vesicles in Figure~\ref{figure2}a are nearly circular, we can
combine~\eqref{eq:stretch},~\eqref{eq:LaplacePressure},
and~\eqref{eq:perm}, to derive
\begin{align}
\label{eq:perm_solution}
L_0(L-L(0)) + L_0^2 \ln\left|\frac{L-L_0}{L(0)-L_0}\right| = -4\pi^2 P K_A t.
\end{align}
All in all, the theoretical time courses for \eqref{eq:perm_solution} are in good
quantitative agreement with the JP data (Figure~\ref{figure2}f).
